### 2023-09-20 paper.tex
Comment version 1.0

@@

However, there has been little research into how decisions of individuals to become entrepreneurs on digital platforms in response to external local incentives lead to systematic difference in eventual performance outcomes across different local contexts. There is no conclusive evidence on how the composition of platform-based entrepreneurs depend on local labor market conditions. This paper aims to bridge these research gaps by examining the following research question. How do local labor market conditions, such as job search frictions and local wages, shape the performance outcomes of platform-based entrepreneurs?

Labor market frictions are are features of the entire human capital system and constitute contextual conditions that allow firms to sustain competitive advantages by appropriating value from strategic human capital investments \citep{mahoney_market_2013,campbell_bridging_2017,starr_strategic_2018}. Location is a systematic source of frictions that operates at the local labor market level, and constrain the ability of workers to move across jobs \citep{dahl_home_2012,kim_entrepreneurial_2021}. Individuals are typically confined to the local wage sector when they look for job opportunities within a region. However, self-employment entails relative flexibility that extends beyond available local employment opportunities, and digital platforms provide a further institutional setting that aims to replace rules and norms with new ones \citep{bonina_digital_2021}. In contrast, individuals become more dependent on specific technology configurations of the platform ecosystem when they join the platform and participate in entrepreneurial activities regardless of their physical location \citep{nambisan_digital_2017,autio_digital_2018,cutolo_platform-dependent_2021}.

To understand how individuals sort into different options at the intersection of platform ecosystems and local labor markets, we draw on the classical theory of labor turnovers that predict workers move across wage jobs and self-employment in response the perceived value of different earnings opportunities \citep{jovanovic_job_1979,jovanovic_matching_1984}. According to the theory, individuals’ actual earnings from platform-based entrepreneurship should increase in their expected income from being employed in a local wage job, when labor markets are fully efficient with many job openings and substantial turnover. However, labor market frictions are departures from efficient markets, and result in disconnect between local wages and entrepreneurial revenues. We hypothesize that such frictions disproportionately affect workers with particularly high human capital value, who are higher ability individuals that tend to outperform in entrepreneurship \citep{elfenbein_small_2010,chatterji_learning_2016}. This entails two predictions: (1) platform-dependent entrepreneurs outperform when they start businesses in situations where local labor market frictions are higher; (2) local labor market frictions suppress the positive correlation between local wages and entrepreneurial revenues.

We formalize these hypotheses with an analytical model, and empirically test them using data on sales volumes and prices of U.S.-based shop owners in a large online marketplace combined with local labor market characteristics from two official data sources – the Job Openings and Labor Turnover Survey (JOLTS) and the Occupational Employment and Wage Statistics (OEWS) database. Based on these publicly available data sources, we construct measures of annual wages and job vacancy rates at the state level, which defines the local labor market for these platform-based entrepreneurs. Empirically, we find that platform-based entrepreneurs who started their businesses during times when vacancy rate in their local labor market was lower ended up earning higher sales revenues. Among those who started their businesses when vacancy rate was higher, the relationship between local wage and entrepreneurial revenue became significantly more positive. Such variation in entrepreneurial performance across different founding year local labor market conditions are likely due to systematic differences in entering sellers' quality. Exploring the potential mechanisms that led to these results, our empirical evidence points to explanations suggesting that different types of sellers (e.g., presence on an external platform such as Instagram, and having a business partner) enter the platform in response to local labor market conditions. These empirical results confirm our theory and hypotheses around local labor market frictions and platform-based entrepreneurial performance.

@@

The paper’s main contribution is to expand our understanding of the determinants of outcomes of platform complementors who create entrepreneurial businesses within the platform ecosystem \citep{cullen_outsourcing_2021}. The phenomenon we study sits at the intersection of two non-overlapping literatures on platform affordances \citep{gawer_bridging_2014,mcintyre_multisided_2021} and local labor market frictions \citep{campbell_rethinking_2012,starr_strategic_2018}, as individuals react to opportunities both within the platform ecosystem and outside the platform in the local market. Our findings confirm that offline economic conditions are non-negligible drivers of the performance of business owners who are complementors in a platform ecosystem.

@@

Digital platforms constitute important contexts in which individuals can start businesses and engage in entrepreneurial activities, through providing products and services to a broad base of both local and non-local customers. While existent literature on local entry conditions and entrepreneurial performance primarily focus on innovation and R\&D intensive ventures \citep{conti2021lowering,bias2022great,hacamo_forced_2022}, how local labor market conditions shape the processes and outcomes of platform-based entrepreneurship remain somewhat underexplored.

@@

On the other hand, platform-based entrepreneurial ecosystems enable less geographically bounded entrepreneurial processes and reduce the costs of communication and barriers to coordination across long geographic distances. Platforms ecosystems enable geographically dispersed individuals to coordinate value creation in diverse location across long distances \citep{cennamo_platform_2013,sussan_digital_2017,jacobides_towards_2018}. The entrepreneurial dynamics within the platform ecosystems are thus less sensitive to local contexts \citep{haveman_spatial_2014,autio_digital_2018}. Entrepreneurs that depend on platforms to conduct businesses are less dependent on local resources. For example, while there are large rural-urban gaps in entrepreneurial activities embedded within local business ecosystems \citep{braesemann_icts_2022}, platform ecosystems provide resources such as access to non-local markets, standardized tools for customer service provision, digital advertising and payment processing \citep{forman_how_2005,koo_platform_2021}. Therefore, relative to localized business ecosystems, platforms enable entrepreneurial activities to move beyond the constraints of geographic boundaries by conducting businesses digitally and reaching remote customers \citep{nambisan_digital_2017,autio_digital_2018}. The entrepreneurial processes on the platform are relatively insulated from demand fluctuations in local product markets, relative to physically embedded businesses that primarily serve local customers.

@@

@@

There is substantial heterogeneity in frictions across urban labor markets in different locations. With a higher state-level vacancy rate, defined by the number of job openings to be filled as a fraction of labor force, it is easier for local individuals to find wage employment. A relatively thinner local labor market has a smaller number of alternative employers, and hence hence the vacancy rate is lower in such markets. Labor market frictions prevent efficient matches between workers and employers from being formed.

@@

\textbf{Performance Outcome.} We measure the performance outcome as the logarithm of estimated seller revenue at a quarterly frequency. Estimated revenue is calculated as the product of quarterly sales volume and average price across listed items at the beginning of the time period. Due to data availability limitations, we do not observe the exact items sold, and hence rely on the average price across all items and total shop-level sales volume to estimate the revenue. In addition, we account for differences in price levels across locations by converting the estimated revenue into real dollars, by diving the nominal revenue by the sate-level price deflator.\footnote{We obtained the regional price deflators from Table 2 of the U.S. Bureau of Economic Analysis report by Figueroa (2022). For each state, the price deflator is expressed as a percentage of 2012 (base-year) national personal consumption expenditure price index.}

@@

\textbf{Local Wage in the Founding Year.} The U.S. Bureau of Labor Statistics makes available the annual employment and wage estimates for different occupations through the Occupational Employment and Wage Statistics (OEWS) program. The raw data (OEWS) provides the annual occupational wages for each state. We use the overall annual wage at the state level as the main measure for local wage, and conduct robustness checks using annual wage for art and design occupations (SOC code $27-1000$). These occupations are most relevant for the entrepreneurs observed in the Etsy sample, because they require skills in art and design that are often involved in craft goods offered by Etsy sellers. We account for differences in price levels across locations by converting the estimated revenue into real dollars, by diving the nominal revenue by the sate-level price deflator.

@@

We find empirical evidence supporting the hypotheses around how local labor market frictions affect performance outcomes of entrepreneurs within a platform ecosystem. Platform-based entrepreneurs systematically outperform when they start the businesses in situations where local labor market frictions are higher. Our empirical results suggest that platform-base entrepreneurial revenues decrease in state-level job vacancy rates. The association between different earnings options – from platform-based entrepreneurship and from local wage employment – becomes more positive when local labor market frictions are reduced, indicated by improved conditions for workers to find wage jobs as vacancy rate increase in the local labor market. Our paper sheds light on the mechanisms that local labor market conditions may affect sorting of skilled individuals into platform-dependent entrepreneurship. When an excess supply of high-quality labor cannot be matched to local jobs due to frictions in the wage sector, they start businesses through a platform ecosystem and outperform in revenue outcome.

>>>>

% some additional comments (Yifan): 1. since the main argument is that the performance difference is due to the entering platform entrepreneurs' quality, is it possible to have some obserables on the quality of the entrepreneurs. Does the entering entrepreneurs during low labour market turnovers have higher quality?

%2. I think this version is much clearer about the terminology. Without excessive mentioning of labur market tightness, it is easier for me to understand. Maybe it is also good to add somewhere in the intro, that the way we observe the labour market friction is to look at e.g., vacancy rate and job turnover etc.

%3. I am not very familiar with the theory development part so please ignore me if I am wrong: I feel many sentences in the theory part express vague meanings or it could be replaced by neat and more precise languages. e.g., Digital platforms constitute important contexts in which individuals can start businesses and engage in entrepreneurial activities... I would probabily just say digital platforms allow/provide an avenue for individuals to start businesses and engage in entrepreneurial activities. And a few paragraphs later, "Digital technologies and platform ecosystems constitute new contexts shaping entrepreneurial decisions and outcomes...", or this "However, local market conditions are push factors that shape the financial incentives for motivating different types of individuals to become entrepreneurs on the platform. " (I'd replace it with more plain sentence) The first sentence meant to give a short and precise summary of this paragraph, but it is not very clear to me what this means. And overall, the literature reviw part of the theory development looks long.

%4. Local labor market condition is main emphasis of this paper. What about the global labor condition or market conditions elsewhere? Does the overall vacancy rate/separation rate in a country/the economy worldwide play a role? What I can imagine is that wage earners/unemployed is probabily attracted by job opportunities in other county/state/country and can migrate to work there (instead of becoming an entrepreneur). Maybe it is good have a few senentences to defend this point.

%5. In the theory model section, I don't think imposing a normal distribution at the begininig is necessary. Because as indicated by the equation (1), any nicely looking distribution with a pdf f(x) and cdf F(x) that is differentiable seem to give the same formula. Plus when we plot the distribution of the revenue, the observable revenue seems not very normally distributed. But the later derivations might change, e.g., if the pdf function is strictly decreasing so that the increase in w would always decrease the observed revenue.

% 5.1. The average entrepreneurial performance of low ability individuals is y^L = E(y_i^L). But is it still normally distributed? or the distribution does not matter?

%6. I also wonder why in the model, the low ability type always choose to be entrepreneurs. In reality, a large number of low ability worker also work on the routine job, earning wages. Is it ok if I make the following argument: there are high ability workers that can choose between entrepreneurs and wage-earners, depending on the market condition. But low ability workers lack talent on either side of the market. They can either choose to be employed (and stay there forever) or become an entrepreneur. And adding this extra component is not going to change the conclusion, as long as there are some low earning individuals who are not affected by market conditions. Intuitively, in the model, the key mechanisim to generate the negative relationship between \mu and observed revenue is that the share of high type entrepreneur relative to the low type entrepreneurs is changing and therefore a rise in \mu (matching rate) makes more high type agents with low realised y_i to be wage-earners. This on the one hand increases the relative share of low type entrepreneurs, driving down mean observed revenues. On the other hand it drives up the average observed revenue among high type entrepreneurs. When the wage rate is high enough, the first effect becomes dominant (because the ave observed revenue among high type entrepreneurs is basically unchanged). Therefore, the argument that "labor market frictions can lead to deviations from efficiency where workers can move freely between wage jobs and platform-dependent entrepreneurship in response to earnings incentives." can hold if the `move more freely' part is more true among the high ability type agents. And I think it is not too difficult to examine this using some micro data.

%7. I feel the argument to reach hypothesis 1 can be imporved. If I understand it correctly, `the platform-dependent entrepreneur generates a higher revenue if the local job vacancy rate was lower' is because  "Labor market frictions prevent efficient matches between workers and employers from being formed", and " such frictions do not directly affect the performance of PDE". Whereas in the model, the important driver seems to be that the self-selection on the high ability typed workers. There won't be any changes in the performance (even a decrease in ave performance due to competition) if any type of workers sort into the PDE in the same proportion.

%8. The conclusion that "Lower labor market frictions are associated with more positive relationship between local wage and revenue from platform-based entrepreneurship." comes from proposition two, where I can see, and you proved that when firctions are lower (\mu are higher), the response of observed performance is higher given fixed changes in wages. But I don't see why equation (2) is positive. If y' increases, could it change from say, -1 to -0.5? I also plotted some matlab graph on the relation between w and revenue, and I dont see this intuition immediately.

@@

% but the sentence is slightly long to read. Maybe it’s good to split it into two?
Little research has been done, however, regarding how the differences in performance outcomes of digital platforms are shaped by the individual decisions to become entrepreneurs, and the decision of these individuals are affected by external local incentives. There is no conclusive evidence on how the composition of platform-based entrepreneurs depend on local labor market conditions. This paper aims to bridge these gaps by examining the following research question: how do local labor market conditions, such as job search frictions and local wages, shape the performance outcomes of platform-based entrepreneurs?

Labor market frictions are important features of the entire human capital system and constitute contextual conditions that allow firms to sustain competitive advantages by appropriating value from strategic human capital investments \citep{mahoney_market_2013,campbell_bridging_2017,starr_strategic_2018}. Location represents a systematic source of frictions that operates at the local labor market level, and constrain the ability of workers to move across jobs \citep{dahl_home_2012,kim_entrepreneurial_2021}. Individuals are typically confined to the local wage sector when they look for job opportunities within a region. However, self-employment entails relative flexibility that extends beyond available local employment opportunities, and digital platforms provide a further institutional setting that aims to replace rules and norms with new ones \citep{bonina_digital_2021}. In contrast, individuals become more dependent on specific technology configurations of the platform ecosystem when they join the platform and participate in entrepreneurial activities regardless of their physical location \citep{nambisan_digital_2017,autio_digital_2018,cutolo_platform-dependent_2021}.

To understand how individuals sort into different options at the intersection of platform ecosystems and local labor markets, we draw on the classical theory of labor turnovers, which predicts that workers move across wage jobs and self-employment in response the perceived value of different earnings opportunities \citep{jovanovic_job_1979,jovanovic_matching_1984}. According to the theory, individuals’ actual earnings from platform-based entrepreneurship should increase in their expected income from being employed in a local wage job, when labor markets are fully efficient with many job openings and substantial turnover. However, labor market frictions are departures from efficient markets, and result in a disconnection between local wages and entrepreneurial revenues. We hypothesize that such frictions disproportionately affect workers with particularly high human capital, who are higher ability individuals that tend to outperform in entrepreneurship \citep{elfenbein_small_2010,chatterji_learning_2016}. This entails two predictions: (1) platform-dependent entrepreneurs outperform when they start businesses in situations where local labor market frictions are higher; (2) local labor market frictions suppress the positive correlation between local wages and entrepreneurial revenues.

We formalize these hypotheses with an analytical model, and empirically test them using data on sales volumes and prices of U.S.-based shop owners in a large online marketplace combined with local labor market characteristics from two official data sources – the Job Openings and Labor Turnover Survey (JOLTS) and the Occupational Employment and Wage Statistics (OEWS) database. Based on these publicly available data sources, we construct measures of annual wages and job vacancy rates at the state level, which defines the local labor market for these platform-based entrepreneurs. Empirically, we find that new platform-based entrepreneurs, who began their businesses in times when the local labor market vacancy rate was lower, achieved higher sales revenues compared to those who started their businesses when the vacancy rate was higher. The latter group exihibits significantly more positive relationship between local wages and entrepreneurial revenue. Such variation in entrepreneurial performance across different local labor market conditions are likely due to systematic differences in entering sellers' quality. Exploring the potential mechanisms that led to these results, our empirical evidence points to explanations suggesting that different types of sellers (e.g., presence on an external platform such as Instagram, and having a business partner) enter the platform in response to local labor market conditions. These empirical results are aligned with our theory and hypotheses around local labor market frictions and platform-based entrepreneurial performance.

@@

% In this paragraph, what do you mean by platform complementors? Are they wage earners or just new entrepreneurs?
The paper’s main contribution is to understand the selection into the platform and what determines the outcomes of platform complementors who create entrepreneurial businesses within the platform ecosystem \citep{cullen_outsourcing_2021}. The phenomenon we study sits at the intersection of two non-overlapping literatures on platform affordances \citep{gawer_bridging_2014,mcintyre_multisided_2021} and local labor market frictions \citep{campbell_rethinking_2012,starr_strategic_2018}, as individuals react to opportunities both within the platform ecosystem and outside the platform in the local market. Our findings confirm that offline economic conditions are non-negligible drivers of the performance of business owners who are complementors in a platform ecosystem.

@@

Digital platforms constitute important contexts in which individuals can start businesses and engage in entrepreneurial activities, through providing products and services to a broad base of both local and non-local customers. While existent literature on local entry conditions and entrepreneurial performance primarily focus on innovation and R\&D intensive ventures \citep{conti2021lowering,bias2022great,hacamo_forced_2022}, how local labor market conditions shape the processes and outcomes of platform-based entrepreneurship remain largely underexplored.

@@

On the other hand, platform-based entrepreneurial ecosystems enable less geographically bounded entrepreneurs and reduce the costs of communication and barriers to coordination across long geographic distances. Platforms ecosystems enable geographically dispersed individuals to coordinate value creation in diverse location across long distances \citep{cennamo_platform_2013,sussan_digital_2017,jacobides_towards_2018}. The entrepreneurial dynamics within the platform ecosystems are thus less sensitive to local contexts \citep{haveman_spatial_2014,autio_digital_2018}. Entrepreneurs that depend on platforms to conduct businesses are less dependent on local resources. For example, while there are large rural-urban gaps in entrepreneurial activities embedded within local business ecosystems \citep{braesemann_icts_2022}, platform ecosystems provide resources such as access to non-local markets, standardized tools for customer service provision, digital advertising and payment processing \citep{forman_how_2005,koo_platform_2021}. Therefore, relative to localized business ecosystems, platforms enable entrepreneurial activities to move beyond the constraints of geographic boundaries by conducting businesses digitally and reaching remote customers \citep{nambisan_digital_2017,autio_digital_2018}. The entrepreneurial processes on the platform are relatively insulated from demand fluctuations in local product markets, relative to physically embedded businesses that primarily serve local customers.

% in the folloiwng sentence, I am not sure if `push factor' (in the area of management) is the correct word to use. Push factors usually refer to factors that encourage people to leave their points of origin and settle somewhere else.

@@

% maybe the last sentence of this paragraph needs to go like: Not only do earnings opportunities with local employers available to the individuals affect the decisions to become entrepreneurs, but also the potential payoffs of entrepreneurship are crucial to shaping transitions into and out of entrepreneurship. (because the first incentive is very obvious, and the second one is what you mentioned in the literature)

@@

There is substantial heterogeneity in frictions across urban labor markets in different locations. With a higher state-level vacancy rate, defined by the number of job openings to be filled as a fraction of labor force, it is easier for local individuals to find wage employment. A relatively thinner local labor market has a smaller number of alternative employers, and hence the vacancy rate is lower in such markets. Labor market frictions prevent efficient matches between workers and employers from being formed.

@@

\textbf{Performance Outcome.} We measure the performance outcome as the logarithm of estimated seller revenue at a quarterly frequency. Estimated revenue is calculated as the product of quarterly sales volume and average price across listed items at the beginning of the time period. Due to data availability limitations, we do not observe the exact items sold, and hence rely on the average price across all items and total shop-level sales volume to estimate the revenue. In addition, we account for differences in price levels across locations by converting the estimated revenue into real dollars, by dividing the nominal revenue by the sate-level price deflator.\footnote{We obtained the regional price deflators from Table 2 of the U.S. Bureau of Economic Analysis report by Figueroa (2022). For each state, the price deflator is expressed as a percentage of 2012 (base-year) national personal consumption expenditure price index.}

@@

\textbf{Local Wage in the Founding Year.} The U.S. Bureau of Labor Statistics makes available the annual employment and wage estimates for different occupations through the Occupational Employment and Wage Statistics (OEWS) program. The raw data (OEWS) provides the annual occupational wages for each state. We use the overall annual wage at the state level as the main measure for local wage, and conduct robustness checks using annual wage for art and design occupations (SOC code $27-1000$). These occupations are most relevant for the entrepreneurs observed in the Etsy sample, because they require skills in art and design that are often involved in craft goods offered by Etsy sellers. We account for differences in price levels across locations by converting the estimated revenue into real dollars, by dividing the nominal revenue by the sate-level price deflator.

@@

We find empirical evidence supporting the hypotheses around how local labor market frictions affect performance outcomes of entrepreneurs within a platform ecosystem. Platform-based entrepreneurs systematically outperform when they start the businesses in situations where local labor market frictions are higher. Our empirical results suggest that platform-based entrepreneurial revenues decrease in state-level job vacancy rates. The association between different earnings options – from platform-based entrepreneurship and from local wage employment – becomes more positive when local labor market frictions are reduced, indicated by improved conditions for workers to find wage jobs as vacancy rate increase in the local labor market. Our paper sheds light on the mechanisms that local labor market conditions may affect sorting of skilled individuals into platform-dependent entrepreneurship. When an excess supply of high-quality labor cannot be matched to local jobs due to frictions in the wage sector, they start businesses through a platform ecosystem and outperform in revenue outcome.
